3.5: Causal Inference and DAGs

### 2019-11-05 static/slides/17-slides.Rmd
Fix gender example in slides 17

1. $gender \rightarrow discr \rightarrow wage$ (Causal path)
2. $gender \rightarrow discr \rightarrow occup \rightarrow wage$ (Causal path)
3. $discr \leftarrow gender \rightarrow occup \rightarrow wage$ (Open back door)
4. $discr \leftarrow gender \rightarrow occup \leftarrow abil \rightarrow wage$ (BLOCKED back door!)
5. $gender \rightarrow discrim \rightarrow occup \leftarrow abil \rightarrow wage$ (BLOCKED back door!)

@@

```{r, echo = F}
dagify(wage~abil+occup+discr,
       discr~gender,
       occup~gender+discr,
       exposure = "disc",
       outcome = "wage") %>% 
  ggdag_parents("wage", stylized = FALSE)+theme_dag_blank()+theme(legend.position = "none")
```

@@

@@

- If leave occupation uncontrolled for: ignore non-discriminatory reasons different genders choose different occupations

@@

- **Can't identify causal effect** by controlling for `occup`!

>>>>

1. $discr \rightarrow wage$ (Causal path)
2. $discr \rightarrow occup \rightarrow wage$ (Causal path)
3. $discr \leftarrow gender \rightarrow occup \rightarrow wage$ (Open back door)
4. $discr \leftarrow gender \rightarrow occup \leftarrow abil \rightarrow wage$ (BLOCKED back door!)
5. $discr \rightarrow occup \leftarrow abil \rightarrow wage$ (BLOCKED back door!)

@@

```{r, echo = F}
dagify(wage~abil+occup+discr,
       discr~gender,
       occup~gender+discr+abil,
       exposure = "disc",
       outcome = "wage") %>% 
  ggdag_parents("wage", stylized = FALSE)+theme_dag_blank()+theme(legend.position = "none")
```

@@

# Collider Example: Gender Discrimination II

@@

@@

- **Can't identify causal effect** controlling for `occup`!
  - *Could* if you could control for `occup` *and* `abil` (but perhaps `abil` is unobserved)
